---
title: "TimeSeries_CrudeoilPrice"
author: "Johnson Adebayo_SCV1007 (Cohort 1)"
date: "3/10/2020"
output: html_document
---

```{r}
oilPrice <- read.csv("BrentOilPrices.csv" )
View(oilPrice) 
```

```{r}
dt_oil = ts(oilPrice,frequency = 365, start = c(1987, 5,20), end = c(2020,2,25))
```


```{r}
class(dt_oil)
```

```{r}
str(dt_oil)
```
```{r}

plot(dt_oil, xlab = "Date", ylab = "Oil Price ($)")
abline(reg=lm(dt_oil~time(dt_oil)))
#abline(reg=lm(dt_oil~time(dt_oil)))
```

```{r}
start(dt_oil)
```

```{r}
end(dt_oil)
```

```{r}
frequency(dt_oil)
```


```{r}
summary(dt_oil)
```

# Plot of the aggregate to show the level and trend
```{r}
plot(aggregate(dt_oil, FUN = mean))
```

```{r}
boxplot(dt_oil~cycle(dt_oil), col=c("red","green","blue","yellow"))
```

```{r}
boxplot(dt_oil, col = "red")
```

```{r}
library(forecast)
seasonplot(dt_oil, 365, col=rainbow(21), year.labels=TRUE, main="Seasonality Plot")
```

```{r}
plot(decompose(dt_oil))
```

```{r}
plot(stl(dt_oil, s.window = "periodic"))
```

Step 2: Stationalize the TS
```{r}
library(tseries)
```

# From the adf.test() below it shows that the dt Time Series (TS) is not stationary because the p-value >0.05, which makes the null hypothesis to be accepted
```{r}
adf.test(dt_oil)
```


```{r}
ndiffs(dt_oil)
```

```{r}
oilprice_diff <- diff(dt_oil, differences = 1)
```


# To check if the oilprice TS is stationary
# from the adf.test above it shows that the TS is stationary
```{r}
adf.test(oilprice_diff)
```

# Plot of Stationarized TS
```{r}
plot(oilprice_diff)
```

# Step 3: To find the parameters

# The auto-correlation function plot below shows that the TS (dt_oil) is not decaying at all because it is not stationary
```{r}
acf(dt_oil)
```

# From the result of acf() on oilprice_diff we obtain the parameters: (p,d,q)=(0,1,1)
# NOTE: my q can be 1,2,3,4,5,6,7,8 etc because the ACF cut off the blue dashed lines at those points which makes MA or q =1,2,3,4,5,6,7,8 etc and AR or p =0
```{r}
acf(oilprice_diff)
```


```{r}
pacf(oilprice_diff)
```


# Step 4: Building the ARIMA model
# Fit ARIMA Model:Let’s fit an ARIMA model
```{r}

# By specifying the order and the seasonality, the stationary has being taken care of.
fit <- arima(dt_oil, order=c(0, 1, 1),seasonal = list(order = c(0, 1, 1), period =4))
```


```{r}
#Price <- oilPrice$Price
#Date <- as.Date(oilPrice$Date)
price <- as.ts(oilPrice$Date,oilPrice$Price)
```

```{r}
str(price)
```

```{r}
plot(oilPrice)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


